---
title: "Age and gender count data"
output: pdf_document
---

# import libraries

```{r}
library(haven) 
library(lubridate)
# library("readxl")
# library(xlsx)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
# library(RColorBrewer)
library(ivreg)
library(broom)
library(generics)
library(here)

# prevent masking
rename = dplyr::rename
select = dplyr::select
summarise = dplyr::summarise
filter = dplyr::filter
```

# import scripts

```{r}

working_dir = here()
script_dir = here('scripts')
plot_dir = here("viz")
csv_dir = here('data')
# paths = c("fear_index.R","conditions_setting.R","data_import.R")
for( path in paths){
  source(file.path(script_dir, path))
}
```

# import dataset and subsetting

```{r}
# get dataset
original <- import_stata(dir=csv_dir, up_to="2021")

# name changing
original =  original %>% 
  rename(age=AdmissionAgeYearepisodeba) 

# select columns
selected_columns = c('eventdate',"DateofRegisteredDeath","age","Sex_num","death","censor","old_home")

# subset dataset
selected = original %>% select(selected_columns)
```

# Impute filtering condition info, group by age and gender

```{r}

# age and gender filtering conditions 
demo_conditions = c("age>= 65 & Sex_num == 1","age >= 65 & Sex_num == 0",
                    "age>=55 & age<65 & Sex_num==1","age>=55 & age<65 & Sex_num==0",
                    "age>=45 & age <55 & Sex_num==1","age>=45 & age <55 & Sex_num==0",
                  "age>=35 & age <45 & Sex_num==1","age>=35 & age <45 & Sex_num==0",
                    "age>=18 & age <35 & Sex_num==1","age>=18 & age <35 & Sex_num==0",
                    "age>=0 & age <18 & Sex_num==1","age>=0 & age <18 & Sex_num==0")
# age and gender filtering conditions in readable form
demo_conditionsTextual = c("65+ & M","65+ & F",
                           "55-64 & M","55-64 & F",
                           "45-54 & M","45-54 & F",
                           "35-44 & M","35-44 & F",
                           "18-34 & M","18-34 & F",
                           "0-17 & M","0-17 & F")
# concat strings in dplyr-readable form
conditions = paste(paste0(demo_conditions,"~\"",demo_conditionsTextual,"\";"),collapse ='')

selected = selected %>% 
  mutate(cond=case_when(
    !!! rlang::parse_exprs(conditions)
  ))
```

# Count data table of age and gender attn

```{r}
attn.age.gender = selected %>% 
  group_by(cond) %>% 
  compare_192021() %>% 
  simplify_yearlycomparison()
attn.age.gender
```

# Count data table of age and gender death

```{r}
death.age.gender = selected %>% 
  filter(death==1) %>%
  group_by(cond) %>% 
  compare_192021() %>% 
  simplify_yearlycomparison()
death.age.gender
```

# Count data of 65+ RCHE attn

```{r}
plus65.attn = selected %>% 
  filter(age>65,old_home==1) %>%
  compare_192021() %>% 
  simplify_yearlycomparison()
plus65.attn
```

# Count data of 65+ RCHE death

```{r}
selected %>% filter(death==1) %>% count(age>65,old_home==1,Year=year(eventdate)) %>% pivot_wider(names_from = "Year", names_glue="n{Year}", values_from = "n") %>% mutate(percent_diff = signif((n2020/n2019-1)*100,3))
```

# Count data of 65+ RCHE among waves

```{r}
daily = selected %>% filter(age>65) %>% count(old_home,eventdate,death) %>% 
  pivot_wider(names_from=death,names_glue="death_{death}",values_from=n,values_fill=0) %>%
  mutate(attn=death_0+death_1) %>%
  rename(death=death_1) %>%
  select(-death_0) %>%
  # mutate year and t, number of calendar days since 1st Jan
  mutate(Year=year(eventdate)-2019,
         t=ifelse(year(eventdate)==2020,
                  eventdate-as.Date("2020-01-01"),
                  eventdate-as.Date("2019-01-01")),
         time_lag = 0) %>%
  filter(eventdate!=as.Date("2020-02-29")) %>%
  drop_na()

# merge with wave settings
daily = daily %>% merge(x=., y=wavePandemicHkIndex.1920, by.x="eventdate", by.y="eventdate")
daily %>% 
  filter(waveFactor != "0") %>% 
  group_by(old_home,waveFactor,Year=year(eventdate)) %>% 
  summarise(attn=sum(attn)) %>% 
  pivot_wider(names_from = "Year", names_glue="n{Year}", values_from = "attn") %>% 
  mutate(attn_percent_diff = signif((n2020/n2019-1)*100,3))
```


# Count data of 65+ 28-day death by DOA/DBA among waves

```{r}
daily = original %>% filter(age>65) %>% count(doa = destination_num==4,eventdate,death) %>% 
  pivot_wider(names_from=death,names_glue="death_{death}",values_from=n,values_fill=0) %>%
  mutate(attn=death_0+death_1) %>%
  rename(death=death_1) %>%
  select(-death_0) %>%
  # mutate year and t, number of calendar days since 1st Jan
  mutate(Year=year(eventdate),
         t=ifelse(year(eventdate)==2020,
                  eventdate-as.Date("2020-01-01"),
                  eventdate-as.Date("2019-01-01")),
         time_lag = 0) %>%
  filter(eventdate!=as.Date("2020-02-29")) %>%
  drop_na()

# merge with wave settings
daily = daily %>% merge(x=., y=wavePandemicHkIndex.1920, by.x="eventdate", by.y="eventdate")

# group by in-wave or out-of-wave periods
daily %>% 
  group_by(doa,During_wave = waveFactor %in% c("1","2","3","4"),Year) %>% 
  summarise(death=sum(death)) %>%
  pivot_wider(names_from = "Year", names_glue="n{Year}", values_from = "death") %>% 
  mutate(death_diff = n2020-n2019,
         death_percent_diff = signif((n2020/n2019-1)*100,3))
```

# Count data of 65+ DOA/DBA death YoY

```{r}
original %>% 
  filter(age>65,death==1) %>% 
  group_by(doa=destination_num==4) %>%
  compare_192021() %>% 
  simplify_yearlycomparison()
```


# Count data of 65+ death YoY

```{r}
original %>% filter(age>65,death==1) %>%
  compare_192021() %>% 
  simplify_yearlycomparison()
```