---
title: "2SLS by age and gender"
output: pdf_document
---

# import libraries

```{r}
library(haven) 
library(lubridate)
library(here)
library(parallel)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(RColorBrewer)
library(ivreg)
library(generics)

# prevent masking
rename = dplyr::rename
select = dplyr::select
summarise = dplyr::summarise
filter = dplyr::filter
here = here::here
```

# import scripts

```{r}
source("./scripts/import.R")
source("./scripts/data_import.r")
```

# import dataset and subsetting

```{r}
# get dataset
stata = mclapply(list("AE_Attendance_transformed_20150101_20210917.csv"), import_csv, mc.cores=40)
stata = stata[[1]]

# select columns
selected_columns = c('eventdate',"DateofRegisteredDeath","age","Sex_num","death","old_home")

# subset dataset
selected = stata %>% 
  select(selected_columns) 

selected = selected %>% 
  mutate(eventdate = as.Date(eventdate)) %>%
  filter(year(eventdate)!=2015) %>%
  # # (optional) discard last 28-day data 
  filter(eventdate<max(eventdate)-28)

max(selected$eventdate)
```

# Impute filtering condition info, group by age and gender

```{r}
# age and gender filtering conditions 
demo_conditions = c("age>= 65 & Sex_num == 1","age >= 65 & Sex_num == 0",
                    "age>=55 & age<65 & Sex_num==1","age>=55 & age<65 & Sex_num==0",
                    "age>=45 & age <55 & Sex_num==1","age>=45 & age <55 & Sex_num==0",
                  "age>=35 & age <45 & Sex_num==1","age>=35 & age <45 & Sex_num==0",
                    "age>=18 & age <35 & Sex_num==1","age>=18 & age <35 & Sex_num==0",
                    "age>=0 & age <18 & Sex_num==1","age>=0 & age <18 & Sex_num==0")
# age and gender filtering conditions in readable form
demo_conditionsTextual = c("65+ & M","65+ & F",
                           "55-64 & M","55-64 & F",
                           "45-54 & M","45-54 & F",
                           "35-44 & M","35-44 & F",
                           "18-34 & M","18-34 & F",
                           "0-17 & M","0-17 & F")
# concat strings in dplyr-readable form
conditions = paste(paste0(demo_conditions,"~\"",demo_conditionsTextual,"\";"),collapse ='')

selected = selected %>% 
  mutate(cond=case_when(
    !!! rlang::parse_exprs(conditions)
  ))


```

# produce daily data table

```{r}
daily = selected %>% group_by(cond) %>% count(eventdate) %>% 
  rename(attn=n) %>%
  # mutate year and t, number of calendar days since 1st Jan
  mutate(Year=year(eventdate)) %>%
  drop_na()

# merge with wave settings
wave.df <- import_waves(twosls = T)
date.df <- import_dates(wave.df=wave.df, 
                        start = min(selected$eventdate), 
                        end=max(selected$eventdate))
# impute after-wave periods
date.df = date.df %>%
  mutate(wave_num = ifelse(is.na(wave_num) & date>min(wave.df$start),"after-wave",wave_num))%>%
  mutate(wave_num = ifelse(date<min(wave.df$start),0,wave_num))
```

# attn reduction per wave, age group and gender (using 2021 data as well)

```{r}
attn.reduction.age = compare_waves(selected, date.df)
attn.reduction = attn.reduction.age %>% 
  # impute gender col
  mutate(gender = ifelse(grepl("M",cond),"M","F"),
           cond = str_sub(cond,1,-5))
```

# Fig2a Attn reduction per wave, age group and gender including 0-17

```{r}
attn_reduction_plot = function(df, conditions){
  attn.reduction = df %>%
    filter(!is.na(wave_num))
    ggplot() +
      geom_line(data=attn.reduction %>% filter(cond %in% conditions, gender=="M",!wave_num %in% c("after-wave","0")),aes(x=wave_num,y=percent_diff,group=1,col=gender)) +
      geom_line(data=attn.reduction %>% filter(cond %in% conditions, gender=="F",!wave_num %in% c("after-wave","0")),aes(x=wave_num,y=percent_diff,group=1,col=gender)) +
      geom_hline(data=attn.reduction %>% filter(cond %in% conditions, gender=="M",wave_num=="after-wave"),aes(yintercept=percent_diff,group=1,col=gender),linetype="dashed") +
      geom_hline(data=attn.reduction %>% filter(cond %in% conditions, gender=="F",wave_num=="after-wave"),aes(yintercept=percent_diff,group=1,col=gender),linetype="dashed") +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            panel.background = element_blank(), axis.line = element_line(colour = "black")) +
      labs(x="Wave Period",y="Seasonally Adjusted Attendance Change %",col="Gender") +
      theme(legend.position = "bottom") +
      facet_wrap(~cond,nrow=1) +
      geom_hline(yintercept = 0)
}
f2a.left = attn_reduction_plot(attn.reduction, c("0-17"))
f2a.left

f2a.middle = attn_reduction_plot(attn.reduction, c("18-34","35-44"))
f2a.middle

f2a.right = attn_reduction_plot(attn.reduction, c("45-54","55-64","65+"))
f2a.right

write_xlsx(attn.reduction %>% filter(wave_num!=0), here::here("results","age_gender_wave_attn.xlsx"))
```

# Fig 2b Bar chart of YoY attn & death change across age groups only

```{r}
results = compare_yearly(selected %>%
                            mutate(year = year(eventdate),
                                   # remove gender from cond
                                   cond = gsub('.{3}$', '', cond)) %>%
                            filter(!is.na(cond)))

cts.overall.age = results[[1]] 
  
cts.percentdiff.overall.age = results[[2]]
```



```{r}
# ///// Fig 2B -----------------------------------------------------------
# Attendance and 28d mortality episode comparison among age groups
fig2b1 =
  ggplot() +
  geom_col(data = filter(cts.overall.age, death == -1, ctrltrt == "ctrl"), aes(x=cond,y=n, fill = ctrltrt), width = 0.5) +
  geom_col(data = filter(cts.overall.age, death == -1, ctrltrt == "trt"), aes(x=cond,y=n, fill = ctrltrt), width = 0.25) +
  geom_text(data = filter(cts.percentdiff.overall.age, death == -1, ctrltrt == "ctrl"),
            aes(x=cond, y=n, label=paste0(percent_diff,"%")),
            position=position_dodge(width=0.9),
            hjust = -.5,
            size=4) +
  labs(x="Age groups",y="Attendance Episode")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position = "none") +
  scale_fill_brewer(palette = "Set2") +
  facet_wrap(~as.factor(year), nrow=2) +
  coord_flip()+
  scale_y_continuous(labels = scales::unit_format(unit = "k", scale = 1e-3,accuracy = 1),
                     limits = c(0, 800000))

fig2b2 = ggplot() +
  geom_col(data = filter(cts.overall.age, death == 1, ctrltrt == "ctrl"), aes(x=cond,y=n, fill = ctrltrt), width = 0.5) +
  geom_col(data = filter(cts.overall.age, death == 1, ctrltrt == "trt"), aes(x=cond,y=n, fill = ctrltrt), width = 0.25) +
  geom_text(data = filter(cts.percentdiff.overall.age, death == 1, ctrltrt == "ctrl"),
            aes(x=cond, y=n, label=paste0(percent_diff,"%")),
            position=position_dodge(width=0.9),
            hjust = -1,
            size=4) +
  labs(x="Age groups",y="28d Death")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position = "none") +
  scale_fill_brewer(palette = "Set2") +
  facet_wrap(~as.factor(year), nrow=2) +
  coord_flip() +
  scale_y_continuous(labels = scales::unit_format(unit = "k", scale = 1e-3,accuracy = 1),
                     limits = c(0, 40000))

f2b = fig2b1 / fig2b2
f2b
write_xlsx(cts.percentdiff.overall.age %>%
            filter(ctrltrt == "ctrl") %>%
            select(-ctrltrt),
          "./results/YoY attn and death change across age groups.xlsx")

```

# fig2c 2SLS

```{r}
daily = selected %>% group_by(cond) %>% count(eventdate,death) %>% 
  pivot_wider(names_from=death,names_glue="death_{death}",values_from=n,values_fill=0) %>%
  mutate(attn=death_0+death_1) %>%
  rename(death=death_1) %>%
  select(-death_0) %>%
  # mutate year and t, number of calendar days since 1st Jan
  mutate(Year=year(eventdate)-2016,
         t=yday(eventdate),
         time_lag = 0) %>%
  drop_na()

# merge with wave settings
wave.df <- import_waves(twosls=T)
date.df <- import_dates(wave.df=wave.df, 
                        start = min(selected$eventdate), 
                        end=max(selected$eventdate))
# impute 2016-2019 wave_num as 0 and after-wave for 2020 and 2021 periods after waves
date.df = date.df %>% 
  mutate(wave_num = ifelse(year<=2019, 0, wave_num)) %>%
  # after-wave periods
  mutate(wave_num = ifelse(is.na(wave_num),"after-wave",wave_num))


daily = daily %>% merge(x=., y=date.df, by.x="eventdate", by.y="date")




daily.w.timelag = daily = daily %>% filter(cond %in% c("65+ & M","65+ & F",
                                              "55-64 & M","55-64 & F",
                                              "45-54 & M","45-54 & F")) %>% 
  # this group_by allows time_lag imputation by condition groups
  group_by(cond)

# introduce time lag of death counts

time_lag_vector = c(1:14)
for(i in time_lag_vector){
   daily.w.timelag = bind_rows(daily.w.timelag, 
                               daily %>% mutate(death = lead(death, i),
                                        time_lag = i))
}
```

# partial f-statistics

```{r}
model_full = lm(attn~wave_num + Year + t + I(t^2) + I(t^3), data = daily %>% filter(time_lag == 0,
                                                       cond == "65+ & M"))
model_reduced = lm(attn~Year + t + I(t^2) + I(t^3), data = daily %>% filter(time_lag == 0,
                                                       cond == "65+ & M"))
result = anova(model_full, model_reduced)

```

# 2SLS

```{r}

ivreg_models_orig = daily.w.timelag %>% drop_na() %>% 
  # introduce time lag
  # 2SLS per condition and time_lag
  group_by(cond,time_lag) %>% 
  do(ivmodels = broom::tidy(ivreg::ivreg(death ~ attn + Year + t + I(t^2) + I(t^3)| wave_num + Year + t + I(t^2) + I(t^3), data = .),
                     conf.int = T, conf.level = 0.95)) %>%
  unnest(ivmodels) 

# save coefficients for prediction graph
ivreg_models_pred = ivreg_models_orig %>%
  filter(time_lag == 0, cond %in% c("65+ & M", "65+ & F")) %>%
  select(cond, term, estimate)
# write.xlsx(as.data.frame(ivreg_models_pred),file="./results/2sls_allcoef_forpredonly.xlsx",row.names=F)

ivreg_models = ivreg_models_orig %>%
  filter(term == "attn")

# (optional) save xlsx of attn coefficient numerics
ivreg_models_xlsx = ivreg_models %>% mutate(estimate = signif(-100* estimate,3),
                                       conf.low = signif(-100 * conf.low, 3),
                                       conf.high = signif(-100* conf.high,3)) %>%
  filter(cond == "65+") %>%
  select(gender, time_lag, estimate, conf.low, conf.high)
write_xlsx(as.data.frame(ivreg_models_xlsx),path="./results/2sls_attn_2016_2021.xlsx")

# impute gender
ivreg_models = ivreg_models %>% mutate(gender = ifelse(grepl("M",cond),"M","F"),
                                       cond = str_sub(cond,1,-5))

f2c = ggplot(data=ivreg_models) +
  geom_line(aes(x=time_lag,y=-estimate * 100,col= gender)) +
  geom_ribbon(aes(x=time_lag,
                  y=-estimate * 100,
                  ymin=-conf.high * 100,
                  ymax=-conf.low * 100,
                  fill= gender),alpha=0.2,show.legend = F) +
  facet_wrap(~cond, nrow=1) +
  labs(x="Time Lag (Days)",
       y="Estimated Excess 28d Death\nPer 100 Reduced ED Visits",
       col="Gender") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        plot.margin = margin(0,0,0,.5, "cm")) +
  theme(legend.position = "bottom") +
  geom_hline(yintercept = 0)

ggsave(plot=f2c,file=file.path(plot_dir, "fig2c_raw.svg"),width=7,height=3.5)
# write_xlsx(as.data.frame(ivreg_models),path="./results/2sls_2021.xlsx")
```



```{r}
design <- 
"ABBDDD
CCCDDD
EEEDDD
"


f2 = ( 
        f2a.left + theme(legend.position = "none", axis.title.x = element_blank()) + 
                labs(y = "Seasonally-Adjusted\nAttendance Change %") 
        + f2a.middle + theme(legend.position = "none", 
                            # axis.text.y = element_blank(),
                            # axis.line.y = element_blank(),
                            # axis.ticks.y = element_blank(),
                            axis.title.y = element_blank()
                            )
        + f2a.right + theme(legend.position = "bottom", 
                            # axis.text.y = element_blank(),
                            # axis.line.y = element_blank(),
                            # axis.ticks.y = element_blank(),
                            axis.title.y = element_blank()
                            )
        + f2b + coord_flip()
        + f2c + theme(legend.position = "bottom")
        + patchwork::plot_layout(design = design)
        + patchwork::plot_annotation(tag_levels = list(c("A", "", "", "B", "","C")))
        )

f2 & theme(text = element_text(size = 12.5))
ggsave(plot=f2 & theme(text = element_text(size = 12.5)), file=file.path(plot_dir,"fig2_raw.svg"), height = 9, width = 11.693)
```



